---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data
   set up working directory
   unzip activity file and read it to R
   ```{r}
   unzip("./activity.zip")
   data <- read.csv("./activity.csv")
   str(data)
   head(data)
   ```
## What is mean total number of steps taken per day?
   Convert date to POSIXct class 
   ```{r}
   library(lubridate)
   data$date <- ymd(data$date)
   str(data)
   ```
   
   Tatal steps daily 
   ```{r}
   datatotal <- aggregate(steps ~ date, data, sum)
   str(datatotal)
   ```
   
   Calculate and report the mean and median
   ```{r}
   meansteps <- mean(datatotal$steps)
   meansteps
   median(datatotal$steps)
   ```
   
   Make a histogram of daily steps 
   ```{r}
   hist(datatotal$steps, breaks = 20, xlab = "Daily Steps", ylab = "Frequency",    main = "Total Steps by Date")
   ```
   
## What is the average daily activity pattern?
   average steps in 5 intervail for all days
   ```{r}
   averagesteps <- aggregate(steps ~ interval, data, mean)
   str(averagesteps)
   plot(steps ~ interval, data = averagesteps, type = "l", xlab = "5 min     
   Interval", ylab = "Mean Daily Steps", main = "Mean Daily Steps by 
   Intervals")
   print
   ```
   
   Find max number of average steps across all days
   ```{r}
   maxsteps <- averagesteps[which.max(averagesteps$steps),1]
   maxsteps
   ```
   
## Imputing missing values
   Count missing values (how many rows with NA)
   ```{r}
   sum(is.na(data$steps))
   ```
   
   replace missing values with mean 
   
   ```{r}
   library(Hmisc)
   newdata <- data
   newdata$steps <- with(newdata, impute(steps, mean))
   str(newdata)
   ```
   Tatal steps daily with new data
   ```{r}
   newdatatotal <- aggregate(steps ~ date, newdata, sum)
   str(newdatatotal)
   ```
   
   Ploting with new data
   ```{r}
   hist(newdatatotal$steps, breaks = 20,  xlab = "Daily Imputed Steps", ylab =  
   "Frequency", main = "Total Imputed Steps by Date")
   ```
   
   Calculate mean and median with new data
   ```{r}
   mean(newdatatotal$steps)
   median(newdatatotal$steps)
   ```
   
   Analysis: The mean and median calculated from the imputed data were almost the same as the  mean and median from              origianl data.  
   
   
## Are there differences in activity patterns between weekdays and weekends?
   Assign weekday and weekend for imputed data
   ```{r}
   newdatawk <- newdata
   newdatawk$typeday <- as.factor(ifelse(weekdays(newdatawk$date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
   str(newdatawk)
   ```
   plotting
   Calculate daily steps 
   ```{r}
   meandailysteps <- aggregate(steps ~ interval + typeday, newdatawk, mean)
   library(lattice)
   xyplot(steps ~ interval | factor(typeday), data = meandailysteps, aspect = 0.8, type = "l")
   ```
   

